|

### 2021-02-02 paper/camera_ready/Manuscript_DS.Rmd
update 2021-02-02 12:15:00

---
author:
  name: |
    [Caspar J. Van Lissa](c.j.vanlissa@uu.nl)$^{a,b}$
    Andreas M. Brandmaier$^{c,d}$
    Loek Brinkman$^{b,e}$
    Anna-Lena Lamprecht$^{b,f}$
    Aaron Peikert$^{c}$
    Marijn E. Struiksma$^{b,g}$
    Barbara Vreede$^{h}$

  affiliation: | 
      $a$: Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics  
      $b$: Open Science Community Utrecht  
      $c$: Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany  
      $d$: Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK  
      $e$: UMC Utrecht  
      $f$: Utrecht University, Department of Information and Computing Sciences  
      $g$: Utrecht University faculty of Humanities, department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS  
      $h$: University Library Utrecht
title: "WORCS: A Workflow for Open Reproducible Code in Science"
short:      "WORKFLOW FOR OPEN SCIENCE"
shorttitle:      "WORKFLOW FOR OPEN SCIENCE"
abstract: |
  Adopting open science
  principles can be challenging and time-intensive, because doing so requires
  substantial conceptual education and training in the use of new tools.
  This paper introduces the Workflow for
  Open Reproducible Code in Science (WORCS): A step-by-step
  procedure that researchers can follow to make a research project open and
  reproducible.
  The purpose of the workflow is to lower the threshold for 
  adoption of open science principles.
  It is based on established best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies.
  To facilitate widespread adoption, the WORCS principles have been implemented
  in the R package `worcs`, which offers an RStudio 
  project template and utility functions for specific workflow steps.
  This paper introduces the conceptual workflow,
  discusses how it meets different 
  standards for open science, and addresses the functionality provided by the
  `R` implementation, `worcs`. 
  This paper is primarily targeted towards
  scholars conducting research projects in R,
  conducting research that involves academic prose, analysis code,
  and (optionally) tabular data.
  However, the workflow is flexible enough to accommodate other scenarios,
  and offers a sensible starting point for customized solutions.
  The source code for the `R` package and manuscript,
  and a list of [user examples of WORCS projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users), are available at 
  https://github.com/cjvanlissa/worcs.
authornote: "blabla"
keywords:
  # at least one keyword must be supplied
  formatted: [open science, reproducibility, r, dynamic document generation, version control]
  plain:     [open science, reproducibility, r, dynamic document generation, version control]
preamble: >
  \usepackage{amsmath}
bibliography      : ["worcs.bib"]  
urlcolor          : blue
output: rmarkdown::html_document
---
<!--
, "worcs_packages.bib"
output: 
  rmarkdown::html_document:
    citation_package: biblatex Note: Citations are always handled by pandoc when format is not PDF
knit: worcs::cite_all
csl: apa.csl
csl: apa.csl
-->

@@

Another important challenge is managing collaborations when only the lead author
uses RMarkdown, and the coauthors use Word. In this case, it is possible to Knit
the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf`
in the manuscript's YAML header to `output: papaja::apa6_docx`. There are some
limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents).
When soliciting feedback from co-authors, ask them to use Track Changes and
comment bubbles in Word. <!-- LB: Overleaf?-->Then, manually incorporate their
changes into the `manuscript.Rmd` file. In most cases, this is the most
user-friendly approach, as most lead authors would review changes by co-authors
anyway. A second approach is to ask collaborators to work in plain text. In this
case, send collaborators the `manuscript.Rmd` file, and ask them to open (and
save) it in Word or Notepad as a plain text file. When they send it back, make
sure any changes to your local file are committed, and then simply overwrite
your local version with their file. In RStudio, select the file in the Git tab,
and click the Diff button to examine what changes the collaborator has made
relative to your last committed version. <!--Finally,
@peikertReproducibleDataAnalysis2019 mention the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] as a
potentially useful way to collaborate with Word users. This package enables
bidirectional conversion between Word and
RMarkdown, with support for tracked changes. However, `redoc` is currently in
suspended development, and for many users, a simple solution (i.e.,
asking collaborators to work in plain text) might be sufficient.-->

>>>>

---
author:
  name: |
    [Caspar J. Van Lissa](c.j.vanlissa@uu.nl ) <a href="https://orcid.org/0000-0002-1825-0097"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{*,a,b}$
    Andreas M. Brandmaier <a href="https://orcid.org/0000-0001-8765-6982"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{c,d}$
    Loek Brinkman <a href="https://orcid.org/0000-0003-3997-1173"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{b,e}$
    Anna-Lena Lamprecht <a href="https://orcid.org/0000-0003-1953-5606"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{b,f}$
    Aaron Peikert <a href="https://orcid.org/0000-0001-7813-818X"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{c}$
    Marijn E. Struiksma <a href="https://orcid.org/0000-0002-1166-1424"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{b,g}$
    Barbara M.I. Vreede <a href="https://orcid.org/0000-0002-5023-4601"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{h,i}$

  affiliation: | 
      $*$: Corresponding author. E-mail: c.j.vanlissa@uu.nl  
      $a$: Department of Methodology & Statistics, Utrecht University, The Netherlands  
      $b$: Open Science Community Utrecht  
      $c$: Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany  
      $d$: Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK  
      $e$: University Medical Center, Utrecht University, The Netherlands  
      $f$: Department of Information and Computing Sciences, Utrecht University, The Netherlands  
      $g$: Utrecht Institute of Linguistics OTS, Utrecht University  
      $h$: University Library, Utrecht University and Netherlands eScience Center  
      $i$: Netherlands eScience Center
title: "WORCS: A Workflow for Open Reproducible Code in Science"
short:      "WORKFLOW FOR OPEN SCIENCE"
shorttitle:      "WORKFLOW FOR OPEN SCIENCE"
abstract: |
  Adopting open science
  principles can be challenging and time-intensive, because doing so requires
  substantial conceptual education and training in the use of new tools.
  This paper introduces the Workflow for
  Open Reproducible Code in Science (WORCS): A step-by-step
  procedure that researchers can follow to make a research project open and
  reproducible.
  The purpose of the workflow is to lower the threshold for 
  adoption of open science principles.
  It is based on established best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies.
  To facilitate widespread adoption, the WORCS principles have been implemented
  in the R package `worcs`, which offers an RStudio 
  project template and utility functions for specific workflow steps.
  This paper introduces the conceptual workflow,
  discusses how it meets different 
  standards for open science, and addresses the functionality provided by the
  `R` implementation, `worcs`. 
  This paper is primarily targeted towards
  scholars conducting research projects in R,
  conducting research that involves academic prose, analysis code,
  and (optionally) tabular data.
  However, the workflow is flexible enough to accommodate other scenarios,
  and offers a sensible starting point for customized solutions.
  The source code for the `R` package and manuscript,
  and a list of [user examples of WORCS projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users), are available at 
  https://github.com/cjvanlissa/worcs.

  Keywords: open science, reproducibility, r, dynamic document generation, version control
authornote: "blabla"
keywords:
  # at least one keyword must be supplied
  formatted: [open science, reproducibility, r, dynamic document generation, version control]
  plain:     [open science, reproducibility, r, dynamic document generation, version control]
preamble: >
  \usepackage{amsmath}
bibliography      : ["worcs.bib"]  
urlcolor          : blue
output            : rmarkdown::html_document
---
<!--
, "worcs_packages.bib"
output: 
  rmarkdown::html_document:
    citation_package: biblatex Note: Citations are always handled by pandoc when format is not PDF
knit: worcs::cite_all
csl: apa.csl
csl: apa.csl
-->

@@

Another important challenge is managing collaborations when only the lead author
uses RMarkdown, and the coauthors use Word. When using the `papaja` package to
write APA-style papers, it is possible to Knit
the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf`
in the manuscript's YAML header to `output: papaja::apa6_docx`. There are some
limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents).
When soliciting feedback from co-authors, ask them to use Track Changes and
comment bubbles in Word. <!-- LB: Overleaf?-->Then, manually incorporate their
changes into the `manuscript.Rmd` file. In most cases, this is the most
user-friendly approach, as most lead authors would review changes by co-authors
anyway. A second approach is to ask collaborators to work in plain text. In this
case, send collaborators the `manuscript.Rmd` file, and ask them to open (and
save) it in Word or Notepad as a plain text file. When they send it back, make
sure any changes to your local file are committed, and then simply overwrite
your local version with their file. In RStudio, select the file in the Git tab,
and click the Diff button to examine what changes the collaborator has made
relative to your last committed version. <!--Finally,
@peikertReproducibleDataAnalysis2019 mention the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] as a
potentially useful way to collaborate with Word users. This package enables
bidirectional conversion between Word and
RMarkdown, with support for tracked changes. However, `redoc` is currently in
suspended development, and for many users, a simple solution (i.e.,
asking collaborators to work in plain text) might be sufficient.-->

### 2021-02-02 paper/camera_ready/preprint.Rmd
update 2021-02-02 12:15:00

---
title: |
    | WORCS: A Workflow for Open Reproducible Code in Science
    | `r paste0("Version: ", worcs:::cran_version(), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`

@@

author:
  - name: "Caspar J. Van Lissa"
    affiliation: "1,2"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
  - name: "Andreas M. Brandmaier"
    affiliation: "3,4"
  - name: "Loek Brinkman"
    affiliation: "2,5"
  - name: "Anna-Lena Lamprecht"
    affiliation: "2,6"
    email: "a.l.lamprecht@uu.nl"
  - name: "Aaron Peikert"
    affiliation: "3"
  - name: "Marijn E. Struiksma"
    affiliation: "2,7"
  - name: "Barbara M.I. Vreede"
    affiliation: "8"

affiliation:
  - id: "1"
    institution: "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id: "2"
    institution: "Open Science Community Utrecht"
  - id: "3"
    institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id: "4"
    institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
  - id: "5"
    institution: "University Medical Center Utrecht"
  - id: "6"
    institution: "Utrecht University, Department of Information and Computing Sciences"
  - id: "7"
    institution: "Utrecht Institute of Linguistics OTS"
  - id: "8"
    institution: "University Library Utrecht"

abstract: |
  Adopting open science
  principles can be challenging and time-intensive, because doing so requires
  substantial conceptual education and training in the use of new tools.
  This paper introduces the Workflow for
  Open Reproducible Code in Science (WORCS): A step-by-step
  procedure that researchers can follow to make a research project open and
  reproducible.
  The purpose of the workflow is to lower the threshold for 
  adoption of open science principles.
  It is based on established best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies.
  To facilitate widespread adoption, the WORCS principles have been implemented
  in the R package `worcs`, which offers an RStudio 
  project template and utility functions for specific workflow steps.
  This paper introduces the conceptual workflow,
  discusses how it meets different 
  standards for open science, and addresses the functionality provided by the
  `R` implementation, `worcs`. 
  This paper is primarily targeted towards
  scholars conducting research projects in R,
  conducting research that involves academic prose, analysis code,
  and (optionally) tabular data.
  However, the workflow is flexible enough to accommodate other scenarios,
  and offers a sensible starting point for customized solutions.
  The source code for the `R` package and manuscript,
  and a list of [user examples of WORCS projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users), are available at 
  https://github.com/cjvanlissa/worcs.

authornote: |
  This is a preprint, not a peer-reviewed publication. Comments and
  contributions are welcome. Please contact the lead author.
  All authors of this paper, except the first, are listed in alphabetical order.
  Author contributions are detailed in the document 'authors.csv' on the project
  repository, at https://github.com/cjvanlissa/worcs.
  The lead author is supported by a NWO Veni grant (NWO grant number
  VI.Veni.191G.090).
  We acknowledge Jeroen Ooms for offering feedback and suggesting the use of the
  R package gert.
keywords          : "open science; reproducibility; r; dynamic document generation; version control"
wordcount         : "8257"

@@

class             : "man"
classoption       : "noextraspace"
output            : papaja::apa6_pdf
urlcolor          : blue
knit              : worcs::cite_all
---

>>>>

---
title: |
    | WORCS: A Workflow for Open Reproducible Code in Science <!--`r paste0("Version: ", worcs:::cran_version(), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`-->

@@

author:
  - name: "Caspar J. Van Lissa \\orcidicon{0000-0002-1825-0097}"
    affiliation: "*,a,b"
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
  - name: "Andreas M. Brandmaier \\orcidicon{0000-0001-8765-6982}"
    affiliation: "c,d"
  - name: "Loek Brinkman \\orcidicon{0000-0003-3997-1173}"
    affiliation: "b,e"
  - name: "Anna-Lena Lamprecht \\orcidicon{0000-0002-5023-4601}"
    affiliation: "b,f"
    email: "a.l.lamprecht@uu.nl"
  - name: "Aaron Peikert \\orcidicon{0000-0001-7813-818X}"
    affiliation: "c"
  - name: "Marijn E. Struiksma \\orcidicon{0000-0002-1166-1424}"
    affiliation: "b,g"
  - name: "Barbara M.I. Vreede \\orcidicon{0000-0002-5023-4601}"
    affiliation: "h,i"

affiliation:
  - id: "a"
    institution: "Department of Methodology & Statistics, Utrecht University, The Netherlands"
  - id: "b"
    institution: "Open Science Community Utrecht"
  - id: "c"
    institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id: "d"
    institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
  - id: "e"
    institution: "University Medical Center, Utrecht University, The Netherlands"
  - id: "f"
    institution: "Department of Information and Computing Sciences, Utrecht University, The Netherlands"
  - id: "g"
    institution: "Utrecht Institute of Linguistics OTS, Utrecht University"
  - id: "h"
    institution: "University Library, Utrecht University and Netherlands eScience Center"
  - id: "i"
    institution: "Netherlands eScience Center"

abstract: |
  Adopting open science
  principles can be challenging, requiring
  conceptual education and training in the use of new tools.
  This paper introduces the Workflow for
  Open Reproducible Code in Science (WORCS): A step-by-step
  procedure that researchers can follow to make a research project open and
  reproducible.
  This workflow intends to lower the threshold for 
  adoption of open science principles.
  It is based on established best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies.
  To facilitate widespread adoption, the WORCS principles have been implemented
  in the R package `worcs`, which offers an RStudio 
  project template and utility functions for specific workflow steps.
  This paper introduces the conceptual workflow,
  discusses how it meets different 
  standards for open science, and addresses the functionality provided by the
  `R` implementation, `worcs`. 
  This paper is primarily targeted towards
  scholars conducting research projects in R,
  conducting research that involves academic prose, analysis code,
  and tabular data.
  However, the workflow is flexible enough to accommodate other scenarios,
  and offers a starting point for customized solutions.
  The source code for the `R` package and manuscript,
  and a list of [examples of WORCS projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users), are available at 
  https://github.com/cjvanlissa/worcs.

authornote: |
  All authors of this paper, except the first, are listed in alphabetical order.
  Author contributions are detailed in the document 'authors.csv' on the project
  repository, at https://github.com/cjvanlissa/worcs.
  The lead author is supported by a NWO Veni grant (NWO grant number
  VI.Veni.191G.090).
  We acknowledge Jeroen Ooms for offering feedback and suggesting the use of the
  R package gert. 
  *Corresponding author. E-mail: c.j.vanlissa@uu.nl
keywords          : "open science; reproducibility; r; dynamic document generation; version control"

@@

class             : "man"
classoption       : "noextraspace"
output:
  papaja::apa6_pdf:
    includes:
      in_header: "preamble.tex"
urlcolor          : blue
knit              : worcs::cite_all
---
